---
title: Example Wind Resource Assessment Using R
author: Stephen Lightfoote
date: '2018-05-30'
always_allow_html: yes
output:
  github_document:
    toc: true
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE,
	fig.align = 'center'
)
```

```{r personal info, message=FALSE, warning=FALSE, include=FALSE}
api_key='YOUR KEY HERE'
name = 'YOUR NAME HERE'
reason = 'wind+resource'
affiliation = 'YOUR AFFILIATION'
email = 'YOUR EMAIL'
mailing_list = 'false'
```

# Example Wind Resource Assessment
The goal here is to illustrate how aspects of typical wind resource assessment and energy capture from meteorological data can be accomplished using open source tools, in this case using R.  Using publicly available data, I'll walk through some of the typical steps taken in site screening, importing, visualizing and analyzing meteorological data with the goal of modeling the annual energy capture of a wind turbine at a given location.  For kicks, as a wink to the NIMBY's out there, let's use my backyard as the example.

## Packages
First, load the R packages used to do the analysis.
```{r, libraries}
library(tidyverse) # r-package ecosystem
library(MASS) # statistical functions
library(knitr) # fancy tables
library(clifro) # windrose plot
library(scales) # percentages
library(elevatr) # elevation data
library(raster) # geospatial
library(leaflet) # mapping
```

## Get Wind Data
Next, we'll download some sample wind resource data from NREL's wind toolkit to see how energetic my location is.  The variable arguments in the paste statement below should be replaced with your own information (see https://developer.nrel.gov/signup/ to get an API key).
```{r, get data}
lat=42.2756887
lon=-71.21161789999996
url<-paste('http://developer.nrel.gov/api/wind-toolkit/wind/wtk_download.csv?api_key=',api_key,'&wkt=POINT(',lon,'%20',lat,
')&attributes=wind_speed,wind_direction,power,temperature,pressure&names=2009&full_name=',name,'&email=',email,'&affiliation=',affiliation,
'&reason=Example',sep='')

#download data as a dataframe
df<-read_csv(url,skip=3,col_types = cols())
# tidy names
names(df)<-gsub("[^[:alnum:]]", "", names(df))
# convert dates to timestamp
df$timestamp_utc=as.POSIXct(paste(df$Year,df$Month,df$Day,df$Hour,df$Minute,'00',sep='-'),format='%Y-%m-%d-%H-%M-%S',tz='UTC')

```

## Exploratory Analysis
Now to the meat of the post, which is to illustrate how to analyze and visualize typical steps for wind resource assessment.

### Data Structure
```{r, data structure}
head(df)
```

### Timeseries
```{r, timeseries}
ggplot(df,aes(timestamp_utc,windspeedat100mms))+geom_line()+theme_minimal()
```

### Monthly Wind Speed Distribution
```{r, monthly winds}
ggplot(df,aes(factor(Month),windspeedat100mms))+
  geom_boxplot()+
  theme_minimal()+
  labs(x='Month')
```

### Wind Rose
```{r, wind rose}
windrose(speed = df$windspeedat100mms,
                 direction = df$winddirectionat100mdeg,
                 n_directions = 12,
                 speed_cuts = seq(0,20,4),
                 ggtheme='minimal',
                 col_pal = 'YlGnBu')
```

### Weibull Fit
```{r, weibull}
weibull_fit<-fitdistr(df$windspeedat100mms,'weibull')
x<-seq(0,20,.01)
weibull_density<-tibble(x,y=dweibull(x = x,shape = weibull_fit$estimate[1],scale = weibull_fit$estimate[2]))
ggplot(df,aes(windspeedat100mms))+
  geom_histogram(aes(y=..density..),bins=30,color='white')+
  geom_line(data=weibull_density,aes(x=x,y=y),color='red')+
  theme_minimal()
```

## Energy Capture
OK, let's model energy capture for this location.

### Power Curve
Let's use the GE 1.5SLE 77m turbine power curve as an example.
```{r, pc}
url<-'http://www.wind-power-program.com/Downloads/Databasepowercurves(May2017).zip'
tmp<-tempfile()
download.file(url,tmp)
unzip(tmp,files = 'Databasepowercurves(May2017)/HAWTs/500kw and above/General Electric/GE 1.5SLE 77m 1.5MW (MG).pow',junkpaths = T)
unlink(tmp)
pc<-read_csv('GE 1.5SLE 77m 1.5MW (MG).pow',
             skip=5,col_names = F,col_types='n',n_max = 30)
pc<-tibble(ws=seq(0.5,29.5,1),kw=pc$X1)
ggplot(pc,aes(ws,kw))+geom_point()+geom_line()+theme_minimal()
```

### Density Adjust Wind Speed
It's important to adjust the raw wind speed by air density as wind power density is a function of air density.
```{r, density adjust}
df<-mutate(df,air_density=(df$surfaceairpressurePa/df$airtemperatureat2mK*287)*.00001,
           dc_ws=windspeedat100mms*(air_density/mean(air_density))^(1/3))
ggplot(df,aes(windspeedat100mms,dc_ws,color=air_density))+
  geom_point()+
  theme_minimal()+
  coord_equal()
```

### Predict Energy
We'll use R's approx function to interpolate the equivalent turbine power for each wind speed in the timeseries.  There's lot's of different methods for this of course.
```{r, energy capture}
df$kw<-approx(pc$ws,pc$kw,df$dc_ws)$y
ggplot(df,aes(kw))+
  geom_density(fill='blue')+
  theme_minimal()
```

## Results

### Aggregates
```{r, aggregate}
monthly<-df %>%
  group_by(Month) %>%
  summarise(hours=n()/12,
            windspeedat100mms=mean(windspeedat100mms),
            mwh=sum(kw,na.rm=T)/(1000*12)) %>%
  mutate(ncf=percent(mwh/(hours*1.5)))
kable(monthly,digits=1,caption='monthly totals',align='c')

#annual
annual<-data.frame(mwh=sum(monthly$mwh),ncf=percent(sum(monthly$mwh/(sum(monthly$hours)*1.5))))
kable(annual,digits=1,caption = 'annual totals',align='c')
```

Nice, a 41.4% NCF for this location is not too shabby.  If only we could convince the landowner :)

### Comparison with NREL Model
NREL's dataset comes with an example power measurement based on a 5MW wind turbine.  Let's see how our simple model stacks up using the GE 1.5SLE.  
```{r, nrel}
ggplot(df,aes(kw*.001,powerMW))+
  geom_point()+
  theme_minimal()+
  labs(x='GE 1.5sle MW',y='NREL 5MW turbine example MW')
```

# Conclusions
Hopefully this was a useful tutorial.  Let me know what you think!
